\Title

### 2023-02-04 Grimaldi-etal-BiolCybernetics.tex
modif legend

Through an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}). First, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogenous synpatic delays. Then, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on. We will study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an efficient solution to the energy/accuracy trade-off. 

@@

Let us now formally define the Heterogeneous Delays Spiking Neural Network (HD-SNN) model, as well as the task to solve and how we will learn to solve it. First, we will define the model for the efficient detection of event-based motifs using a feed-forward layer of spiking neurons with heterogeneous delays. Our HD-SNN model will be tested on a computationally relevant task of detecting motion detection from an event-based sensory input. By analogy with biological conditions, we will design a paradigm simulating eye movements where an input image is translated by a parameterized saccadic trajectory. Solving this task is, for example, useful in real-life situations to compensate for eye, head or body movements and to provide robust image categorization by stabilizing the image on the retina. The task is thus to guess the motion as accurately as possible (\seeFig{motion_task}) and we will describe a semi-supervised learning mechanism for weights and delays. To this end, we will extend the HD-SNN model to be efficient for the resolution of the task by including spatio-temporal convolutional kernels. We will finally describe its implementation on conventional computers. 

@@

    {\bf - Motion detection task.} {\bf (Left)}~We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view such as to reproduce the effect of displacing the eye. To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here an example path of the trajectory as a green line ($200$ time steps). {\bf (Right)}~Snapshot of the synthesized event stream at different time steps. The dynamics of the sub-image as a function of time generates a naturalistic movie which may be transformed into an event-based representation. Mimicking the retina, this representation codes for proportional and pixel-wise increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events. On the bottom left corner of the images, the translation vector are plotted in red among the different possible classes of movements. Note the change of direction between the second and third frames.}

>>>>

Through an ecological motion detection task, we study the emergence of spatio-temporal spiking motifs when training a single layer of spiking neurons on a semi-supervised classification task (\seeFig{motion_task}). First, we define the ecological cognitive task the model has to solve with the different datasets it will be tested on. Then, we develop a theoretically defined Heterogeneous Delays in Spiking Neural Network (HD-SNN) model capable of learning heterogenous synaptic delays. We will study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an efficient solution to the energy/accuracy trade-off. 

@@

@@

    {\bf - Motion detection task.} {\bf (Left)}~We use large-scale natural images ($512\times512$) in which an aperture ($128\times128$) extracts a sub-image image in the axis of view. To mimic the effect of a saccadic-like eye movement, the axis of view moves following a step-like random walk, and we show here an example path of the trajectory as a green line ($200$ time steps). {\bf (Right)}~Snapshot of the synthesized event stream at different time steps. The dynamics of the sub-image as a function of time generates a naturalistic movie which may be transformed into an event-based representation. Mimicking the retina, this representation codes for proportional and pixel-wise increments or decrements of the luminance in the image, respectively ON (in red) and OFF (in blue) events. On the bottom left corner of the images, the translation vector are plotted in red among the different possible classes of movements. Note the change of direction between the second and third frames.}
